Modal Abstractions for Virtualizing Memory Addresses

### 2024-07-10 examples.tex
+

Looking at the update from Specfication Lines 5-10 to 10-14 in Figure \ref{walkpgdir}, we see that the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \textsf{pte-ptsto} (Specfication Line 13 in Figure \ref{walkpgdir}) needed for executing the load instruction (Line 15 in Figure \ref{fig:calltopteinitialize}) to obtain the value of a page entry whose physical address (\textsf{entry}) kept in the register $\texttt{r8}$. This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.

@@

  \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast
   \exists \textsf{freshaddr, entry} \textsf{entry\_pfn} entry freshaddr = paddr \ast freshaddr \mapsto_{\textsf{p}} \textsf{w0\_64}  \leftrightarrow \mathsf{val} = \mathsf{w0\_64}\urcorner \; \ast\\
   \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad   \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast 
% \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    
\underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
;;void pte_initialize(pte_t **table) {
...
;;set up the stack    
;;allocate a full zeroed page for 512 8-byte entries 
;;pte_t *local = kalloc();
callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
mov    rax,-0x10[rbp] 
;;*table = local;
mov    -0x10[rbp],rax
mov    -0x8[rbp],rdi
mov    rax,[rdi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{id}} \textsf{\_}\ast \forall_{i\in\textsf{0..511}} \textsf{nextpaddr + i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
;;clean up the stack
..
\end{lstlisting}
\vspace{-1em}
\caption{Allocating a physical page }
\label{pteinitializespec}
\end{figure}
Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 34 and 38 in Figure \ref{fig:calltopteinitialize}. There are two important aspects on relating a freshly allocated page-table physical address to the identity-mapping which we focus in more detail separately in the specification of \textsf{pte\_initialize} in Figure \ref{fig:pteinitializespec}. First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource. Second, we see the identity mapping for the entry frame ($\textsf{entry\_val}_{\textsf{pfn}}$), and all the knowledge on the existence of the identity mappings for both the frame and the page entries indexed from the frame (Line 38 in Figure \ref{fig:calltopteinitialize}) identity mappings exists, which are not part of the invariant yet, which could occur once the presence bit of the entry is set (Line 42 in Figure \ref{fig:calltopteinitialize}). Presence bit guard in the extended invariant shown in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} \todo{we need to say something about the improtance of being present and bein inside the invariant.} allows us to preserve the \textit{soundness} of the table-traversal it eliminates all the entries (and tables) that are being under mutation for any other access other than mutation itself.
%%%%deprecated 
\begin{comment}
once the presence bit of the entry for which the invariant  the entry}first not appearing in the identity map (Specification Line 11 in Figure \ref{fig:pteinitializespec}), then as inserted into the identity-map with a value \textsf{Z} ($\Xi \cup \{\mathsf{nextpaddr}\}$) together with the fragmental capability  -- $\ulcorner \Xi \; !!\; \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- at Lines 13 and 14 in Figure \ref{fig:pteinitializespec} . The value type \textsf{Z}, unlike the value type \textsf{V} denoting the random value in Line 8 in Figure \ref{fig:calltopteinitialize}, denotes \textsf{zeroed} value.

@@

@@

 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$

>>>>

Looking at the update from Specfication Lines 3 to 15 in Figure \ref{walkpgdir}, we see that the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \textsf{pte-ptsto} (Specfication Line 15 in Figure \ref{walkpgdir}) needed for executing the load instruction (Line 15 in Figure \ref{fig:calltopteinitialize}) to obtain the value of a page entry whose physical address (\textsf{entry}) kept in the register $\texttt{r8}$. This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.

@@

  \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast  \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\
   \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \paddr + \textsf{i * 8} \mapsto_{\textsf{id,qfacc}} \textsf{mvalidc} \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast \\

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
;;void pte_initialize(pte_t **table) {
...
;;set up the stack    
;;allocate a full zeroed page for 512 8-byte entries 
;;pte_t *local = kalloc();
callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
mov    rax,-0x10[rbp] 
;;*table = local;
mov    -0x10[rbp],rax
mov    -0x8[rbp],rdi
mov    rax,[rdi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{id}} \textsf{\_}\ast \forall_{i\in\textsf{0..511}} \textsf{nextpaddr + i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
;;clean up the stack
...
\end{lstlisting}
\vspace{-1em}
\caption{Allocating a physical page }
\label{pteinitializespec}
\end{figure}
Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 34 and 38 in Figure \ref{fig:calltopteinitialize}. There are two important aspects on relating a freshly allocated page-table physical address to the identity-mapping which we focus in more detail separately in the specification of \textsf{pte\_initialize} in Figure \ref{fig:pteinitializespec}. First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource. Second, we see the identity mapping for the entry frame ($\textsf{entry\_val}_{\textsf{pfn}}$), and all the knowledge on the existence of the identity mappings for both the frame and the page entries indexed from the frame (Line 38 in Figure \ref{fig:calltopteinitialize}) identity mappings exists, which are not part of the invariant yet, which could occur once the presence bit of the entry is set (Line 42 in Figure \ref{fig:calltopteinitialize}). Presence bit guard in the extended invariant shown in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} \todo{we need to say something about the improtance of being present and bein inside the invariant.} allows us to preserve the \textit{soundness} of the table-traversal it eliminates all the entries (and tables) that are being under mutation for any other access other than mutation itself.
%%%%deprecated 
\begin{comment}
once the presence bit of the entry for which the invariant  the entry}first not appearing in the identity map (Specification Line 11 in Figure \ref{fig:pteinitializespec}), then as inserted into the identity-map with a value \textsf{Z} ($\Xi \cup \{\mathsf{nextpaddr}\}$) together with the fragmental capability  -- $\ulcorner \Xi \; !!\; \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- at Lines 13 and 14 in Figure \ref{fig:pteinitializespec} . The value type \textsf{Z}, unlike the value type \textsf{V} denoting the random value in Line 8 in Figure \ref{fig:calltopteinitialize}, denotes \textsf{zeroed} value.

@@

$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_}}_{\rtv}$

@@

 $\specline{\textsf{P} \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
